---
title: "p8105_hw5_tw2918"
output: github_document
date: "2023-11-08"
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Probelm 1
### Describe the raw data.

**Import and Clean Dataset**

* update the victim's first and last name to proper case
* replace unknown to missing value NA
* update victim_age to numeric variable

```{r}
homicide_df = read_csv("./data/homicide-data.csv") |>
  janitor::clean_names()|>
  mutate (
    victim_last = str_to_title(victim_last),
    victim_first = str_to_title(victim_first),
    across(c(uid, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, disposition), ~na_if(., "Unknown")),
    victim_age = as.numeric(victim_age)
  )
```

**Describe the raw data**

* The datasaet contains data on homicides in 50 large U.S. cities from  2007 to 2017. The dataset contains `r nrow(homicide_df)` total observations and `r ncol(homicide_df)` variables. 
* The variable records the id number of the cases (`uid`), the report date (`reported_date`),cities and state of the cases (`city` and `state`), the detailed case location (`lat` and `lon`) and the case status on whether it is open and closed (`disposition`). It also recoreds victim's name (`victim_last` and `victim_first`), race (`victim_race`), age (`victim_age`) and gender(`victim_sex`). 

### Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
homicide_df = homicide_df |>
  mutate(city_state = paste(city, state, sep = ", "))

homicide_disposition = homicide_df|>
  group_by(city_state)|>
  summarise(
    total_homicides = sum(disposition %in% c("Closed by arrest","Closed without arrest", "Open/No arrest"), na.rm = TRUE),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"), na.rm = TRUE)
    )

print(homicide_disposition)
```

### For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
homicide_disposition_baltimore = homicide_disposition |>
  filter(city_state == "Baltimore, MD")

unsovled = homicide_disposition_baltimore |> 
  pull(unsolved_homicides)

total = homicide_disposition_baltimore |> 
  pull(total_homicides)

prop_test_baltimore = prop.test(unsovled, total)|>
  broom::tidy()|>
  select(estimate, conf.low, conf.high)

print(prop_test_baltimore)
```

### run prop.test for each of the cities in the dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
homicide_city = homicide_disposition |>
  mutate(
    prop_test_city = map2(unsolved_homicides, total_homicides, ~prop.test(.x,.y)),
    tidy_test_city = map(prop_test_city, broom::tidy)
  )|>
  unnest(tidy_test_city)|>
  select(city_state, estimate, conf.low, conf.high)
  
print(homicide_city)
```

### Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r}
homicide_city |>
  mutate(city_state = reorder(city_state, estimate))|>
  ggplot(aes(x=city_state, y =estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))+
  labs(
    title = "Proportion Estimate and Confidance Interval of Unsolved Homicides In US Cities ",
    x = "Cities and States",
    y = "Proportion Estimate",
  ) +
  theme(axis.text.x =element_text(angle = 90, hjust = 1))
  
```

## Problem 2

### Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time

* Start with a dataframe containing all file names using `list.files` function 
* Iterate over file names and read in data for each subject using `purrr::map` and saving the result as a new variable in the dataframe
* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
longit_study_df = 
  tibble(
    file_name = list.files("./data/p2", full.names = TRUE))|> 
    # using `full.names = True` to use the file path as the value instead of file name
  mutate(
    data = purrr::map (file_name, ~ read_csv(.x, show_col_types = FALSE)),
    subject_id = str_extract(file_name, "\\d{2}"), 
    #extract 2 digits numbers from the file_name
    arm = if_else(str_detect(file_name, "con"), "control", "experimental"),
    #if detect "con" in `file_name`, make the value "control", else make the value "experimental" 
    subject_id = fct_relevel(subject_id, "01", "02", "03", "04", "05", "06", "07", "08", "09", "10"),
  )|>
  select(subject_id, arm, data)|>
  unnest(data)|>
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observation"
  ) |>
  mutate(week = as.numeric(week))
```

### Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r}
longit_study_df |>
  ggplot(aes(x = week, y = observation,)) +
  geom_point(aes(color = subject_id)) +
  geom_line(aes(color = subject_id)) +
  geom_smooth(aes(color = arm), method = "loess", se = FALSE) +
  labs(
    title = "Obsservation on Each Subject Over Time ",
    x = "Time in Week",
    y = "Observation Data",
    color = "ID and Arm"
    ) 
```

```{r}
longit_study_df |>
  ggplot(aes(x = week, y = observation, color = subject_id)) +
  geom_point() +
  geom_line() +
  facet_grid(~arm)
  labs(
    title = "Obsservation on Each Subject Over Time ",
    x = "Time in Week",
    y = "Observation Data",
    color = "ID and Arm"
    ) 
```

*comment on differences between groups*

* The starting observation value in week 1 is similar and mixed between the two arms. The observation value increases for the experimental arm over time and is higher than the control arm starting in week 2. The observation value for control group stayed stable aover time from week1 to week 8. 

## Problem 3

### set design elements and function

```{r}
n = 30
sigma = 5
mu = 0
num_simulations = 5000
alpha = 0.05
set.seed(1)


sim_ttest = function (n, mu, sigma) {
  
  sample = rnorm (n, mean = mu, sd = sigma)
  t_test = t.test(sample, mu=0)
  result = broom::tidy(t_test) |>
    select(estimate, p.value)
  
}

sim_results = map_dfr(1:num_simulations, ~ sim_ttest(n, mu, sigma))

print(sim_results)

```


### Repeat the above for  μ={1,2,3,4,5,6}

```{r}
mu_value = 1:6
set.seed(1)

sim_ttest = function (n, mu, sigma) {
  
  sample = rnorm (n, mean = mu, sd = sigma)
  t_test = t.test(sample, mu=0)
  result = broom::tidy(t_test) |>
    select(estimate, p.value)
  
}

sim_mu_and_ttest = function(mu) {
  map_dfr(1:num_simulations, function(x) sim_ttest(n=30, mu=mu, sigma = 5))
}

sim_results_all = map_dfr(mu_value, sim_mu_and_ttest, .id = "mu")|>
  janitor::clean_names()|>
  mutate(mu = as.numeric(mu))

print(sim_results_all)

```

### Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.


```{r}
sim_results_all |>
  group_by(mu)|>
  summarize (power = mean(p_value < 0.05))|>
  ggplot(aes(x = mu, y = power))+
  geom_point() +
  geom_line() +
  labs(
    title = "Power Over True Value of μ ",
    x = "True Value of μ",
    y = "Power of the Test"
    ) 
```

*Association between effect size and power*

* The plot shows a positive association between effect size and power that as the true value of μ increases, the power of the test also increases, indicating larger effect sizes are easier to detect. When the effect size is small (values of μ close to 0), the test has less power, which means it is less likely to reject the null hypothesis even if it is false. As the effect size grows, the test becomes more likely to reject the null hypothesis.
* The plot shows that the the slope or the increase in power is getting flatter, as the true value μ gets large, suggesting that once the effect size is sufficiently large, further increases in the effect size do not substantially increase power, likely because the power is already close to 1. At this point, the test will almost always correctly reject the null hypothesis.

### Make a plot showing the average estimate of μ̂ on the y axis and the true value of μ on the x axis, and the average estimate of μ̂ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis

```{r}
sim_results_all |>
  group_by(mu)|>
  summarize (
    average_estimate_all = mean(estimate),
    average_estimate_rejected = mean(estimate[p_value < alpha])
    )|>
  pivot_longer(
    average_estimate_all: average_estimate_rejected,
    names_to = "average_estimate_group",
    names_prefix = "average_estimate_",
    values_to = "average_estimate_value"
  )|>
  ggplot(aes(x = mu, y = average_estimate_value, color = average_estimate_group))+
  geom_point() +
  geom_line() +
  labs(
    title = "Average Estimate of μ Hat Over True Value of μ ",
    x = "True Value of μ",
    y = "Average Estimate of μ Hat"
    ) 
```

*Result*

* From the plot, it appears that the average estimate of when the null is rejected (yellow line) is consistently above the line for all tests (purple line), and both are increasing with the true value of μ.
* The sample average of μ̂ across tests for which the null is rejected does not equal to the true value of μ when the true value of μ is small. As the true value of μ increase, the sample average of μ̂ across tests for which the null is reject is getting closer to the true value and approximately equal to the true value when the true value is equal and above 4. 
* This is because, when the true value of μ is small,there's magnitude error, which is rejecting null hypothesis base on random chance or sampling variability. Therefore the the average estimate from these samples is biased upwards. And when the true value of μ is large, the power of the test becomes high, and the sample estimates of μ are more consistently close to the true μ. 
