---
title: "p8105_hw5_tw2918"
output: github_document
date: "2023-11-08"
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Probelm 1
### Describe the raw data.

**Import and Clean Dataset**

* update the victim's first and last name to proper case
* replace unknown to missing value NA
* update victim_age to numeric variable

```{r}
homicide_df = read_csv("./data/homicide-data.csv") |>
  janitor::clean_names()|>
  mutate (
    victim_last = str_to_title(victim_last),
    victim_first = str_to_title(victim_first),
    across(c(uid, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, disposition), ~na_if(., "Unknown")),
    victim_age = as.numeric(victim_age)
  )
```

**Describe the raw data**

* The datasaet contains data on homicides in 50 large U.S. cities from  2007 to 2017. The dataset contains `r nrow(homicide_df)` total observations and `r ncol(homicide_df)` variables. 
* The variable records the id number of the cases (`uid`), the report date (`reported_date`),cities and state of the cases (`city` and `state`), the detailed case location (`lat` and `lon`) and the case status on whether it is open and closed (`disposition`). It also recoreds victim's name (`victim_last` and `victim_first`), race (`victim_race`), age (`victim_age`) and gender(`victim_sex`). 

### Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
homicide_df = homicide_df |>
  mutate(city_state = paste(city, state, sep = ", "))

homicide_disposition = homicide_df|>
  group_by(city_state)|>
  summarise(
    total_homicides = sum(disposition %in% c("Closed by arrest","Closed without arrest", "Open/No arrest"), na.rm = TRUE),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"), na.rm = TRUE)
    )

print(homicide_disposition)
```

### For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
homicide_disposition_baltimore = homicide_disposition |>
  filter(city_state == "Baltimore, MD")

unsovled = homicide_disposition_baltimore |> 
  pull(unsolved_homicides)

total = homicide_disposition_baltimore |> 
  pull(total_homicides)

prop_test_baltimore = prop.test(unsovled, total)|>
  broom::tidy()|>
  select(estimate, conf.low, conf.high)

print(prop_test_baltimore)
```

### run prop.test for each of the cities in the dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
homicide_city = homicide_disposition |>
  mutate(
    prop_test_city = map2(unsolved_homicides, total_homicides, ~prop.test(.x,.y)),
    tidy_test_city = map(prop_test_city, broom::tidy)
  )|>
  unnest(tidy_test_city)|>
  select(city_state, estimate, conf.low, conf.high)
  
print(homicide_city)
```

