---
title: "p8105_hw5_tw2918"
output: github_document
date: "2023-11-08"
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Probelm 1
### Describe the raw data.

**Import and Clean Dataset**

* update the victim's first and last name to proper case
* replace unknown to missing value NA
* update victim_age to numeric variable

```{r}
homicide_df = read_csv("./data/homicide-data.csv") |>
  janitor::clean_names()|>
  mutate (
    victim_last = str_to_title(victim_last),
    victim_first = str_to_title(victim_first),
    across(c(uid, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, disposition), ~na_if(., "Unknown")),
    victim_age = as.numeric(victim_age)
  )
```

**Describe the raw data**

* The datasaet contains data on homicides in 50 large U.S. cities from  2007 to 2017. The dataset contains `r nrow(homicide_df)` total observations and `r ncol(homicide_df)` variables. 
* The variable records the id number of the cases (`uid`), the report date (`reported_date`),cities and state of the cases (`city` and `state`), the detailed case location (`lat` and `lon`) and the case status on whether it is open and closed (`disposition`). It also recoreds victim's name (`victim_last` and `victim_first`), race (`victim_race`), age (`victim_age`) and gender(`victim_sex`). 

### Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
homicide_df = homicide_df |>
  mutate(city_state = paste(city, state, sep = ", "))

homicide_disposition = homicide_df|>
  group_by(city_state)|>
  summarise(
    total_homicides = sum(disposition %in% c("Closed by arrest","Closed without arrest", "Open/No arrest"), na.rm = TRUE),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"), na.rm = TRUE)
    )

print(homicide_disposition)
```

### For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
homicide_disposition_baltimore = homicide_disposition |>
  filter(city_state == "Baltimore, MD")

unsovled = homicide_disposition_baltimore |> 
  pull(unsolved_homicides)

total = homicide_disposition_baltimore |> 
  pull(total_homicides)

prop_test_baltimore = prop.test(unsovled, total)|>
  broom::tidy()|>
  select(estimate, conf.low, conf.high)

print(prop_test_baltimore)
```

### run prop.test for each of the cities in the dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
homicide_city = homicide_disposition |>
  mutate(
    prop_test_city = map2(unsolved_homicides, total_homicides, ~prop.test(.x,.y)),
    tidy_test_city = map(prop_test_city, broom::tidy)
  )|>
  unnest(tidy_test_city)|>
  select(city_state, estimate, conf.low, conf.high)
  
print(homicide_city)
```

### Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r}
homicide_city |>
  mutate(city_state = reorder(city_state, estimate))|>
  ggplot(aes(x=city_state, y =estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))+
  labs(
    title = "Proportion Estimate and Confidance Interval of Unsolved Homicides In US Cities ",
    x = "Cities and States",
    y = "Proportion Estimate",
  ) +
  theme(axis.text.x =element_text(angle = 90, hjust = 1))
  
```

## Problem 2

### Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time

* Start with a dataframe containing all file names using `list.files` function 
* Iterate over file names and read in data for each subject using `purrr::map` and saving the result as a new variable in the dataframe
* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
longit_study_df = 
  tibble(
    file_name = list.files("./data/p2", full.names = TRUE))|> 
    # using `full.names = True` to use the file path as the value instead of file name
  mutate(
    data = purrr::map (file_name, ~ read_csv(.x, show_col_types = FALSE)),
    subject_id = str_extract(file_name, "\\d{2}"), 
    #extract 2 digits numbers from the file_name
    arm = if_else(str_detect(file_name, "con"), "control", "experimental"),
    #if detect "con" in `file_name`, make the value "control", else make the value "experimental" 
    subject_id = fct_relevel(subject_id, "01", "02", "03", "04", "05", "06", "07", "08", "09", "10"),
  )|>
  select(subject_id, arm, data)|>
  unnest(data)|>
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observation"
  ) |>
  mutate(week = as.numeric(week))
```

### Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r}
longit_study_df |>
  ggplot(aes(x = week, y = observation,)) +
  geom_point(aes(color = subject_id)) +
  geom_line(aes(color = subject_id)) +
  geom_smooth(aes(color = arm), method = "loess", se = FALSE) +
  labs(
    title = "Obsservation on Each Subject Over Time ",
    x = "Time in Week",
    y = "Observation Data",
    color = "ID and Arm"
    ) 
```

```{r}
longit_study_df |>
  ggplot(aes(x = week, y = observation, color = arm)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Obsservation on Each Subject Over Time ",
    x = "Time in Week",
    y = "Observation Data",
    color = "ID and Arm"
    ) 
```

*comment on differences between groups*

* The starting observation value in week 1 is similar and mixed between the two arms. The observation value increases for the experimental arm over time and is higher than the control arm starting in week 2. The observation value for control group stayed stable aover time from week1 to week 8. 

## Problem 3

### set design elements and function

```{r}
n = 30
sigma = 5
mu = 0
num_simulations = 5000
alpha = 0.05
set.seed(1)


sim_ttest = function (n=30, mu=0, sigma=5) {
  
  sample = rnorm (n, mean = mu, sd = sigma)
  t_test = t.test(sample, mu=mu)
  result = broom::tidy(t_test) |>
    select(estimate, p.value)
  
}

sim_results = map_dfr(1:num_simulations, ~ sim_ttest())

print(sim_results)

```


### Repeat the above for  μ={1,2,3,4,5,6}

```{r}
n = 30
sigma = 5
mu_value = 1:6
num_simulations = 5000
alpha = 0.05
set.seed(1)


sim_ttest = function (n=30, mu=1, sigma=5) {
  
  sample = rnorm (n, mean = mu, sd = sigma)
  t_test = t.test(sample, mu=mu)
  result = broom::tidy(t_test) |>
    select(estimate, p.value)
  
}

sim_mu_and_ttest = function(mu) {
  map_dfr(1:num_simulations, function(x) sim_ttest())
}

sim_results_all = map_dfr(mu_value, sim_mu_and_ttest, .id = "mu")

print(sim_results_all)

```

